---
title: "Gene Set Enrichment and Pathways Analysis"
author: "Mark Dunning"
date: "6 March 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading the data

```{r}
deTable <- read.csv("t47d_Treatment_DEA_Prog-vs-Control_all.csv")
deTable
```

## Initial Galaxy Workflow

- Load csv file
- Convert to tabular
- Remove header line by selecting last lines of file
- Cut to columns 1 to 7 and 12 (Gene Symbol)

### Produce GSEA input file in Galaxy

- Sort by column 5
- Cut columns 1 and 8

The corresponding code in R:-

Now make a ranked-list table for input into GSEA

```{r}

deTable %>% arrange(desc(stat)) %>% 
  dplyr:::select(symbol, stat) %>% 
  filter(!is.na(symbol)) %>% 
  write.table("t47d_Treatment_DEA_Prog-vs-Control_all_for_GSEA.rnk",row.names=FALSE,sep="\t",quote=FALSE)

```



## Gene-set Enrichment analysis with `goseq`

The `goseq` package can be used to find over- and under-represented gene sets. The analysis requires a vector of all genes that were analysed with a logical `TRUE`/`FALSE` value to indicate whether that gene was differentially-expressed in the contrast of interest.

## Preparing a file for Galaxy

```{r}
## If we have been using annotation packages, select might be overwritten to be select from the AnnotationDBI package rather than dplyr. Using dplyr:::select we explicitly make sure the correct function is used.

library(dplyr)
deTable %>% mutate(padj = ifelse(is.na(padj), 1, padj)) %>% 
  mutate(DE = ifelse(padj < 0.05, TRUE, FALSE)) %>% 
  dplyr:::select(X, DE) %>% 
  write.table("t47d_Treatment_DEA_Prog-vs-Control_all_for_Galaxy.tsv",sep="\t",row.names=FALSE)
```


### Workflow in Galaxy
Galaxy Workflow to annotate whether each gene is DE

- Compute expression `c7 <0.05`
- Cut columns 1 and 9


### Preparing data for goseq

We will select genes with adjusted p-value `<0.05`. 
```{r}
genes <- rep(FALSE,nrow(deTable))
genes[which(deTable$padj < 0.05)] <- TRUE
table(genes)
names(genes) <- deTable$X 
```


```{r}
library(goseq)
pwf <- nullp(genes, "hg19", id = "ensGene")
goRes <- goseq(pwf, "hg19","ensGene")
goRes
```

